DisjunctiveProgramming.jl: Generalized Disjunctive Programming Models and Algorithms for JuMP

### 2023-10-14 paper/paper.tex
update paper

\begin{align*}
    \bigvee_{i \in I} Y_i \bigvee_{j \in J} (\neg Y_j) & \ \ \text{becomes} \ \ \sum_{i \in I} y_i + \sum_{j \in J} (1-y_j) \geq 1 \\
\end{align*}

@@

\subsection{Solution Techniques}

@@

\subsection{Example}
To illustrate the syntax in \verb|DisjunctiveProgramming.jl| (Version 0.3.3), consider the simple superstructure optimization problem for the chemical process given in Figure \ref{fig:superstruct_opt_diagram}. In this problem a chemical plant with two candidate reactor technologies ($R_1$ and $R_2$) must be designed. If the second reactor technology is chosen, a separation system must also be installed, for which two separation technologies ($S_1$ and $S_2$) are available. The GDP model seeks to maximize the product flow ($F_7$), while discounting for reactor ($C_R$) and separator ($C_S$) installation costs as given in \eqref{eq:example_obj}, subject to the nested disjunction in \eqref{eq:example_gdp} and the global mass balances in \eqref{eq:example_global}
- \eqref{eq:example_global1}. The system variables are the flows on each stream $i$ ($F_i$) and the installation costs, with their respective bounds given in \eqref{eq:example_var1} - \eqref{eq:example_var3}. The fixed cost and process yield parameters are given by $\gamma$ and $\beta$, respectively.

@@

# add global constraints to model
@constraints(m,
    begin
        F[1] == F[2] + F[3]
        F[7] == F[5] + F[6]
    end
)
\end{lstlisting}
    \item Define the inner (nested) disjunction for the separation technologies in the superstructure using the \verb|@disjunction| macro.   
\begin{lstlisting}[language = Julia]
# define constraints in left YS disjunct
@constraint(m, F[5] == β[:S1]*F[4], DisjunctConstraint(YS[1]))
@constraint(m, CS == γ[:S1], DisjunctConstraint(YS[1]))
# define constraints in right YS disjunct
@constraint(m, F[5] == β[:S2]*F[4], DisjunctConstraint(YS[2]))
@constraint(m, CS == γ[:S2], DisjunctConstraint(YS[2]))
# define disjunction (specify parent disjunct)
@disjunction(m, YS, DisjunctConstraint(YR[2]))
\end{lstlisting}
    \item Define the outer disjunctions.
\begin{lstlisting}[language = Julia]
# define constraints in left YR disjunct
@constraint(m, F[6] == β[:R1]*F[2], DisjunctConstraint(YR[1]))
@constraint(m, [i = 3:5], F[i] == 0, DisjunctConstraint(YR[1]))
@constraint(m, CR == γ[:R1], DisjunctConstraint(YR[1]))
@constraint(m, CS == 0, DisjunctConstraint(YR[1]))
# define constraints in right YR disjunct
@constraint(m, F[6] == β[:R2]*F[3], DisjunctConstraint(YR[2]))
@constraint(m, CR == γ[:R2], DisjunctConstraint(YR[2]))  
# define disjunction
@disjunction(m, YR)    
\end{lstlisting}
    \item Add the selection logical constraints using the \verb|choose!| function. The first constraint enforces that only one reactor is selected (i.e., $Y_{R_1} \ \underline{\vee} \ Y_{R_2}$). The second constraint enforces that the separation system be defined only if the second reactor ($R_2$) is selected. This constraint is equivalent to the proposition $Y_{R_2} \Leftrightarrow Y_{S_1} \ \underline{\vee} \ Y_{S_2}$.
\begin{lstlisting}[language = Julia]
@constraint(m, YR in Exactly(1))
@constraint(m, YS in Exactly(YR[2]))
\end{lstlisting}
    \item Add the objective function and optimize. 
\begin{lstlisting}[language = Julia]
@objective(m, Max, F[7] - CS - CR)
optimize!(m)
\end{lstlisting}
\end{enumerate}

\section{Future Work}
The next steps for the \verb|DisjunctivePrograming.jl| package rely on extending \verb|JuMP.jl| further to allow creating GDP models that are not necessarily reformulated at model creation. Such models will allow using the different GDP solution strategies, such as direct reformulation to MI(N)LP, disjunctive branch and bound, logic-based outer approximation, hybrid cutting planes, and basic steps. The updated package will leverage existing JuMP extension infrastructure and make it possible to define indexing notation for disjunctions, a new Boolean variable type, and a new disjunciton constraint type. These improvements are expected to make the package more usable, flexible, and performant for advanced applications of GDP in JuMP.

>>>>

\begin{align*}
    \left(\bigvee_{i \in I} Y_i\right) \vee \left(\bigvee_{j \in J} \neg Y_j\right) & \ \ \text{becomes} \ \ \sum_{i \in I} y_i + \sum_{j \in J} (1-y_j) \geq 1 \\
\end{align*}

@@

\subsection{Other Solution Techniques} \label{other_techniques}

@@

\subsection{Example}
To illustrate the syntax in \verb|DisjunctiveProgramming.jl| (Version 0.4.0), consider the simple superstructure optimization problem for the chemical process given in Figure \ref{fig:superstruct_opt_diagram}. In this problem a chemical plant with two candidate reactor technologies ($R_1$ and $R_2$) must be designed. If the second reactor technology is chosen, a separation system must also be installed, for which two separation technologies ($S_1$ and $S_2$) are available. The GDP model seeks to maximize the product flow ($F_7$), while discounting for reactor ($C_R$) and separator ($C_S$) installation costs as given in \eqref{eq:example_obj}, subject to the nested disjunction in \eqref{eq:example_gdp} and the global mass balances in \eqref{eq:example_global}
- \eqref{eq:example_global1}. The system variables are the flows on each stream $i$ ($F_i$) and the installation costs, with their respective bounds given in \eqref{eq:example_var1} - \eqref{eq:example_var3}. The fixed cost and process yield parameters are given by $\gamma$ and $\beta$, respectively.

@@

# add global constraints to model
@constraints(m,
    begin
        F[1] == F[2] + F[3]
        F[7] == F[5] + F[6]
    end
)
\end{lstlisting}
    \item Define the inner (nested) disjunction for the separation technologies in the superstructure using the \verb|@disjunction| macro.   
\begin{lstlisting}[language = Julia]
# define constraints in left YS disjunct
YS1_disjunct = DisjunctConstraint(YS[1])
@constraint(m, F[5] == β[:S1]*F[4], YS1_disjunct)
@constraint(m, CS == γ[:S1], YS1_disjunct)
# define constraints in right YS disjunct
YS2_disjunct = DisjunctConstraint(YS[2])
@constraint(m, F[5] == β[:S2]*F[4], YS2_disjunct)
@constraint(m, CS == γ[:S2], YS2_disjunct)
# define disjunction (specify parent disjunct)
@disjunction(m, YS, YS2_disjunct)
\end{lstlisting}
    \item Define the outer disjunctions.
\begin{lstlisting}[language = Julia]
# define constraints in left YR disjunct
YR1_disjunct = DisjunctConstraint(YR[1])
@constraint(m, F[6] == β[:R1]*F[2], YR1_disjunct)
@constraint(m, [i = 3:5], F[i] == 0, YR1_disjunct)
@constraint(m, CR == γ[:R1], YR1_disjunct)
@constraint(m, CS == 0, YR1_disjunct)
# define constraints in right YR disjunct
YR2_disjunct = DisjunctConstraint(YR[2])
@constraint(m, F[6] == β[:R2]*F[3], YR2_disjunct)
@constraint(m, CR == γ[:R2], YR2_disjunct)  
# define disjunction
@disjunction(m, YR)    
\end{lstlisting}
    \item Add the selection logical constraints using the \verb|choose!| function. The first constraint enforces that only one reactor is selected (i.e., $Y_{R_1} \ \underline{\vee} \ Y_{R_2}$). The second constraint enforces that the separation system be defined only if the second reactor ($R_2$) is selected. This constraint is equivalent to the proposition $Y_{R_2} \Leftrightarrow Y_{S_1} \ \underline{\vee} \ Y_{S_2}$.
\begin{lstlisting}[language = Julia]
@constraint(m, YR in Exactly(1))
@constraint(m, YS in Exactly(YR[2]))
\end{lstlisting}
    \item Add the objective function and optimize. 
\begin{lstlisting}[language = Julia]
@objective(m, Max, F[7] - CS - CR)
optimize!(m)
\end{lstlisting}
\end{enumerate}

\section{Future Work}
Since the package is currently limited to reformulating GDP models into Mixed-Integer Programming JuMP models, future work includes developing logic-based solvers to allow optimizing GDP models directly. This will include GDP solvers that implement the other solution techniques described in Section \ref{other_techniques}. Future work also involves extending the list of suported reformulation techniques to include the \textit{P-Split} \cite{kronqvist2022p} and \textit{True-False} \cite{agarwal2015novel} reformulations.
